Polka-dotted Stars: a Hierarchical Model for Mapping Stellar Surfaces Using Occultation Light Curves and the Case of TOI-3884

### 2025-02-27 src/tex/ms.tex
changes

\newcommand{\sbu}{Department of Physics and Astronomy, Stony Brook University, Stony Brook NY 11794, USA}
\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute, New York NY 10010, USA}

@@

% Title
\title{Polka-dotted Stars: a Hierarchical Model for Mapping Stellar Surfaces Using Occultation Light Curves}

@@

% Main body with filler text
\section{Introduction}
\label{sec:intro}
The quest to unravel the mysteries of distant planetary systems has led to a fast-evolving field of research, where exoplanet studies are at the forefront of 
astronomical investigations. Characterizing the surfaces of exoplanet-host stars is crucial for understanding the environments and evolution of their planetary 
systems. For example, the Sun's magnetic field drives the star's activity cycles, the heating of its outer atmosphere and the solar wind,
which affects the Earth's magnetosphere and biosphere \citep{Babcock1961,Charbonneau2014}. 
Therefore, constructing a catalogue of the magnetic properties of other stars similar to the Sun would support the quest towards understanding 
the population of detected exoplanets and the development of a highly generalizable theory on stellar magnetic dynamo.

The reverse statement is also true. The exisiting popultion of exoplanets along with their orbital parameters can help characterize their host-stars
in terms of their surface magnetic activity. Stellar surface features like spots and flares impact observables in planet-detection research including photometric 
variability, spectral line profiles, and radial velocity shifts. However, directly resolving stellar surfaces remains challenging.

While high-resolution spectroscopy provides the most robust method of determining stellar parameters, obtaining such data for large samples remains 
observationally expensive. Some recent works offered invaluable insights into the topographical characteristics of stars and their magnetic 
activity, responsible for the observed variatians on their surfaces. For example, the influence of star spots can be discerned through photometric 
data acquired from Kepler or TESS. \cite{Luger2021b} investigated the star spots features on the stellar surfaces using Gaussian Process (GPs) 
with a \emph{physically interpretable} kernel. They showed that it is hard to put constrains on star spot properties if you only have a light curve for 
one star, it is rather usefule though to perform the same analysis for an ensemble of stars. However, when a star has an exoplanet in orbit, 
the spot features become observable in the transit data, providing valuable constraints on the properties of these spots 
(whether they are dark or bright, their sizes, and their latitudes).

Data-driven approaches like Gaussian processes (GPs) offer pathways to map the surfaces of exoplanet hosts through measurable disk-integrated signals. 
In this work, we develop a Gaussian process framework to model the surfaces of exoplanet-host stars based on time-series photometry. We utilize 
expressive kernel function from \cite{Luger2021b}. We first use this model on a simulated data to test the model, then we apply the model to an active K-dwarf
HAT-P-11.

The planet-hosting star HAT-P-11 has been extensively studied for its star spots. HAT-P-11 is orbited by a Neptune-sized exoplanet, HAT-P-11b, 
which was discovered in 2008 via the transit method \citep{Bakos2010,Deming2011}. The transiting nature of HAT-P-11b has allowed for detailed 
characterization of the star spot activity on HAT-P-11, as star spots that are occulted during the planet's transit cause detectable brightening in 
the transit light curve \citep{Morris2017}.

%
% \section{The Data}
% The data is collected by the \emph{Kepler} mission and we extracted it using
% \texttt{lightkurve}, a Python package for Kepler and TESS data analysis \citep{lightkurve}.
% %
% \begin{figure}[ht!]
%     \script{TransitFitsWithStarry.py}
%     \begin{centering}
%         \includegraphics[width=\linewidth]{figures/TransitFitsWithStarry.pdf}
%         \caption{
%             Transit fits with starry -- no GPs.
%         }
%         \label{fig:TransitFitsStarry}
%     \end{centering}
% \end{figure}
%
\section{Hierarchical Bayesian model}
In this section, we describe our Gaussian Process (GP) model and the likelihood calculation process used to estimate the model parameters.
The light curves obtained from telescopic observations often provide insufficient information for comprehensive photometric studies of stellar 
magnetic fields, yet they also yield an overabundance of data for simple forward modeling. To address the first issue, we focus on capturing the essence 
of the variability caused by the magnetic dynamo rather than solving it in great detail. The second challenge can be overcome by employing Gaussian Processes and 
marginalizing over spherical harmonic maps. 
We approximate the likelihood as a multidimensional Gaussian distribution characterized by a mean $$\pmb{\mu}$$ and covariance $$\pmb{B}$$ (for a comprehensive 
derivation of the Gaussian Process, refer to \cite{Luger2021b}):
%
\begin{align}
    \mathcal{L}\left(i_\star, P_\star, \mathbf{u}, \pmb{\theta}_\bullet\right) \sim
    \mathcal{N}\Big(
    \pmb{\mu}\left(i_\star, P_\star, \mathbf{u}, \pmb{\theta}_\bullet\right),
    \,
    \pmb{B}\left(i_\star, P_\star, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \Big).
\end{align}
%
To infer the physical properties of the starposts, we want to obtain an underlying distribution of the physics conditioned on the provided data, 
which is known as the posterior distribution of the model.
% \subsection{Model for non-evolvng surfaces} \label{sec:nonevolmodel} 
We want to solve for a large set of parameters that includes the GP hyperparameters, the star's and the planet's orbital parameters, respectively. 
\begin{linenomath}\begin{align}
    \label{eq:largetheta}
    \pmb{\Theta}
     & =
    \left(
    \theta_\bullet
    \,\,\,
    \theta_\star
    \,\,\,
    \theta_p
    \right)^\top
    \quad,
\end{align}\end{linenomath}

@@

We represent the GP hyperparameters as \emph{physically interesting} set of parameters $\pmb{\theta}_\bullet$ \citep{Luger2021b}:
%
\begin{linenomath}\begin{align}
        \label{eq:thetaspot}
        \pmb{\theta}_\bullet
         & =
        \left(
        \mathbb{n}
        \,\,\,
        \mathbb{c}
        \,\,\,
        \mathbb{a}
        \,\,\,
        \mathbb{b}
        \,\,\,
        \mathbb{r}
        \right)^\top
        \quad,
    \end{align}\end{linenomath}
%
where $\mathbb{n}$ is the number of starspots, $\mathbb{c}$ is their contrast (defined as the intensity difference between the spot and the 
background intensity, as a fraction of the background intensity), and $\mathbb{r}$ is the radius
of the spots. $\mathbb{a}$ and $\mathbb{b}$ are the normalized parameters of the Beta distribution in $\cos\phi$, which is the probability density function (PDF) 
for the latitude $\phi$ of the spots. These parameters have a one-to-one correspondence to the mode $\mu_\phi$ and standard deviation $\sigma_\phi$ of the 
distribution in $\phi$, allowing for a concise representation of the latitudinal distribution characteristics.
The Beta distribution in $\cos\phi$ which has hyperparameters $\alpha$ and $\beta$, and the PDF given by
%
\begin{align}
    \label{eq:cosphi-pdf}
    p \big(\cos\phi \, \big| \, \alpha, \beta \big)
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    (\cos\phi)^{\alpha - 1}
    (1 - \cos\phi)^{\beta - 1}
    \quad,
\end{align}
%
where $\Gamma$ is the Gamma function. The $\alpha$ and $\beta$ are derived from the normalized parameters 
%
\begin{align}
    \label{eq:beta2gauss}
    \alpha & = \exp\left({K_{00} + (\ln\frac{1}{2})\mathbb{a}}\right)
    \nonumber                                                 \\
    \beta  & = \exp\left({\ln\frac{1}{2} + (10 - \ln\frac{1}{2})\mathbb{b}}\right)
    \quad,
\end{align}
%
with inverse transform
%
\begin{align}
    \label{eq:gauss2beta}
    \mathbb{a} & \equiv \frac{\ln\alpha}{\ln\frac{1}{2}}
    \nonumber                                             \\[0.5em]
    \mathbb{b} & \equiv \frac{\ln\beta - \ln\frac{1}{2}}{10 -\ln\frac{1}{2}}
\end{align}
%
The parameters $\mathbb{a}$ and $\mathbb{b}$ are both constrained to values between 0 and 1, making them convenient for sampling during the inference process. 
However, $a$ and $b$ do not have a straightforward relationship with physically meaningful quantities. In many situations, it is preferable to parametrize 
the latitude distribution using two parameters: $\mu_\phi$, which controls the central latitude, and $\sigma_\phi$, which governs the dispersion of the 
spots' latitudes. Here, both the mean $\mu_\phi$ and variance $\sigma^2_\phi$ can be derived from the Beta distribution as
%
\begin{align}
    \label{eq:mean_beta}
    \mu_\phi
     & =
    \dfrac{\alpha}{\alpha+\beta}
    \quad,
\end{align}
%
\begin{align}
    \label{eq:var_beta}
    \sigma^2_\phi
     & =
    \dfrac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}
    \quad,
\end{align}
%

We assume that the prior over $\mathbb{f}_{true}$ follows a multivariate Gaussian distribution, with a mean vector of zeros and a covariance 
matrix $\pmb{\Sigma}$. We use the quasi-periodic kernel to define the covariance matrix $\pmb{\Sigma}$, which is defined by \texttt{StarryProcess}.
We assume that the observations $\mathbb{f}_{obs}$ are corrupted by additive Gaussian noise, such that:
\begin{equation}
    \mathbb{f}_{obs} = \mathbb{f}_{true} + \epsilon
\end{equation}

@@

\subsubsection{The issue of the normalization of light curves and units}
Before going on discussing the experiments produced for this paper, we first need to remind the reader of a subtlety of the flux normalization when we are given
the raw light curves from telescopes. The problem and the ways to tackle it were described in \cite{Luger2021a} and \cite{Luger2021b}.

@@

In our experiments, we noticed that dividing by a constant factor \emph{and} adding an additive baseline messes up the units due to the 
intricate transition between the spherical harmonics and flux unit bases. To address this issue, we opt for an alternative approach, wherein we 
impose a constraint upon the constant terms within the map, which correspond to a featureless star. Specifically, we achieve this by setting the prior distribution 
over these constant terms to be entirely uniform, thereby setting the first row of the precision matrix to zero.

@@

\subsection{Model for evolving surfaces}
The model described in Section \ref{sec:nonevolmodel} would not work on stars whose surface features evolve repidly in time. Some active stars, particularly 
Sun-like stars, exhibit surface evolution on timescales as short as a single rotation period. For example, starspots can appear and dissipate within days or weeks, 
while chromospheric plages and coronal loops can flare up and decay on even shorter timescales. Such rapid evolution of surface structures violates 
the fundamental assumption of a static photosphere underlying the model described in Section \ref{sec:nonevolmodel}. 
Accounting for these time-dependent surface phenomena is crucial for accurately interpreting observations of active stars, as they can introduce significant 
time-varying distortions in the photometric light curves.

>>>>

\newcommand{\sbu}{Department of Physics and Astronomy, Stony Brook University, Stony Brook NY 11794, USA}
\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute, New York NY 10010, USA}
\renewcommand{\thefigure}{\arabic{figure}}

@@

% Title
\title{Polka-dotted Stars: a Hierarchical Model for Mapping Stellar Surfaces Using Occultation Light Curves and the Case of TOI-3884}

@@

% Main body with filler text
\section{Introduction}
\label{sec:intro}
The quest to unravel the mysteries of distant planetary systems has led to a fast-evolving field of research, where exoplanet studies are at the forefront of 
astronomical investigations. The detection of atmospheres on Earth analogs using transmission spectroscopy is a challenging task due to the confounding 
effects of stellar contamination (see, e.g., \cite{Ducrot2018, Morris2018}). Stellar magnetic activity, particularly in the form of starspots, can significantly 
influence the observed transmission spectra, making it difficult to disentangle planetary atmospheric signals from those originating from the host star. 
Starspots exhibit variations in both time and wavelength, mainly due to stellar rotation. As a result, the identification of starspots has the potential to 
complicate the analysis and interpretation of exoplanet observations.

Therefore, characterizing the behavior of stellar surface features that arise due to magnetic activity like starspots and flares is instrumental for improving 
planet detection efficiency in photometric, spectral, and radial velocity studies. Understanding starspots is fundamental and imminent for the ongoing 
space mission JWST, which will characterize tens to hundreds of exoplanet atmospheres around stars with strong variability. In fact, recent JWST 
results (\cite{Lim2023}) on the atmosphere of TRAPPIST-1b showed that accurate measurements have so far been substantially stymied by stellar contamination, 
and we need more accurate models to account for it. To fully capitalize on the unprecedented capabilities of JWST and accurately interpret the obtained 
transmission spectra, it is crucial to develop a comprehensive understanding of stellar magnetic activity and its impact on the observed signals.

Moreover, characterizing the surfaces of exoplanet-host stars is crucial for understanding the environments and evolution of their planetary 
systems. For example, the Sun's magnetic field drives the star's activity cycles, the heating of its outer atmosphere and the solar wind,
which affects the Earth's magnetosphere and biosphere \citep{Babcock1961,Charbonneau2014}. 
Therefore, constructing a catalogue of the magnetic properties of other stars similar to the Sun would support the quest towards understanding 
the population of detected exoplanets and the development of a highly generalizable theory on stellar magnetic dynamo.

The reverse statement is also true. The exisiting popultion of exoplanets along with their orbital parameters can help characterize their host-stars
in terms of their surface magnetic activity. Stellar surface features like spots and flares impact observables in planet-detection research including photometric 
variability, spectral line profiles, and radial velocity shifts. However, directly resolving stellar surfaces remains challenging. 
While high-resolution spectroscopy provides the most robust method of determining stellar parameters, obtaining such data for large samples remains 
observationally expensive, but photometric observations can come to rescue. Some recent works offered invaluable insights into the topographical 
characteristics of stars and their magnetic activity, responsible for the observed variatians on their surfaces. For example, the influence of starspots 
can be seen through photometric data acquired from Kepler or TESS. \cite{Luger2021b} investigated the star spots features on the stellar surfaces using 
Gaussian Process (GPs) with a \emph{physically interpretable} kernel. They argue that the complex degeneracies make it impossible to determine the 
precise configuration of starspots and other features on a star's surface solely based on its rotational light curve, but we don't have to be 
concerned with the specific characteristics of an individual spot, as such information would have limited practical utility. Instead, we often consider 
(either explicitly or implicitly) the properties of a star spot as a sample drawn from an underlying parent distribution that governs aspects like the average 
and dispersion of spot radii. The parameters that control this distribution are the ones that can be predicted through physical principles, 
making them the primary focus of our interest.

\cite{Luger2021b} also showed that it is hard to put constrains on starspot properties if you only have a light curve for 
one star, it is rather usefule though to perform the same analysis for an ensemble of stars. However, photometric light curves capturing planetary transits encode 
a wealth of information about the stellar surface and the starspots present on it. Transits, during which the planet occults different regions of 
the stellar disk, provide a unique opportunity to probe the physical characteristics of starspots. Exoplanet host stars are \textit{uniquely} suited 
for photometric studies of starspot properties: spot features of observed exoplanet host stars in transit can provide valuable 
constraints on spot properties (see Figure \ref{fig:cartoon}) -- brightness, size, and latitude -- while fitting a single light curve for a star 
with no observed transit introduces too many degeneracies.

In this work, we utilize two powerful computational tools, \texttt{starry} and \texttt{StarryProcess}, to model the light curves of spotted stars and 
their transiting exoplanets. We then extend these models to create a comprehensive framework that combines both stellar variability and planetary transits.

The \texttt{starry} code, developed by \cite{Luger2019}, provides an efficient method for computing light curves of stars and planets. 
It uses a spherical harmonic framework to represent the surface brightness distribution of celestial bodies, allowing for fast and accurate 
calculations of rotational light curves and occultations. \texttt{starry} is particularly useful for modeling the effects of stellar spots 
and planetary transits, as it can handle complex surface brightness patterns and compute their resulting light curves with high precision.

Building upon \texttt{starry}, \cite{Luger2021b} introduced \texttt{StarryProcess}, a probabilistic model for stellar variability. \texttt{StarryProcess} 
treats the distribution of spots on a stellar surface as a Gaussian process in spherical harmonic space. This approach allows for the efficient 
modeling of stellar variability over long time scales, accounting for the evolution and rotational modulation of starspots. \texttt{StarryProcess} 
provides a flexible framework for inferring stellar and starspot properties.

Our model combines the capabilities of both \texttt{starry} and \texttt{StarryProcess}, integrating them with a transit model to 
create a comprehensive framework for analyzing light curves of spotted stars with transiting exoplanets. By leveraging the spherical harmonic 
representation from \texttt{starry} and the probabilistic approach of \texttt{StarryProcess}, we can simultaneously model the stellar variability, 
spot distribution, and planetary transits. This combined approach allows us to account for the effects of stellar spots on transit light curves, 
including spot-crossing events, while also constraining the overall stellar variability pattern. This integrated model enables us to derive more 
accurate and precise constraints on both stellar and planetary parameters. It accounts for the interplay between stellar activity and planetary transits, 
allowing us to disentangle the effects of spots from true transit features. Furthermore, by modeling multiple transits simultaneously with the 
underlying stellar variability, we can better constrain parameters such as stellar inclination, obliquity, and spot distribution, 
while also improving our estimates of planetary properties.

\begin{figure*}[hbt!]
    \includegraphics[width=\textwidth]{figures/star-planet-cartoon.png}
        \caption{Light curves from the kinds of stars shown on the left panel (these light curves were calculated analytically 
        using \texttt{starry} \citep{Luger2019}): (a) the light curve shows a flat flux if the star has no surface features and no planet orbiting it; 
        (b) a transit in the light curve of the star that hosts an exoplanet; (c) if the star has one orbiting planet and one starspot under the planet's 
        trajectory then it is seen as a bump in a transit signal; (d) the light curve becomes complicated when the star has multiple spots.}
        \label{fig:cartoon}
    \end{figure*}

The light curves obtained from telescopic observations often provide insufficient information for comprehensive photometric studies of stellar 
magnetic fields, yet they also yield an overabundance of data for simple forward modeling. To address the first issue, we focus on capturing the essence 
of the variability caused by the magnetic dynamo rather than solving it in great detail (as per the argument in \cite{Luger2021b}). The second challenge can 
be overcome by marginalizing over (integrating out) the properties of each individual spot on a star to infer the characteristics of the overall spot distribution. 
This could be accomplished by modeling the properties of each spot and generating 
the corresponding light curves with a model like \texttt{starry}, which does exactly that by constructing a disk-integrating signal 
with spherical harmonics. Then one can use a posterior sampling scheme, such as Markov Chain Monte Carlo (MCMC), 
while incorporating hyperparameters that control the distribution of spot properties across the entire population. The marginal posteriors for 
the hyperparameters would ultimately provide the information we seek. Bayesian modeling provides a principled framework for incorporating prior 
knowledge and uncertainties into the inference process, allowing us to robustly estimate the hyperparameters governing the distribution of star spot 
properties while naturally accounting for the complex degeneracies and multi-modality inherent in the problem.
However, in practice, the degeneracies and often extreme multi-modality of the distributions of individual spot properties would render this approach 
quite challenging and computationally expensive. That's why we harness the elegant machinery like Gaussian processes (GPs) to perform 
this marginalization on our behalf, simplifying the process and reducing the computational burden.

In Figure \ref{fig:nullspace}, we present an analysis similar to Figure 4 in \cite{Luger2021b}, but with the addition of a planetary transit. 
This figure illustrates the decomposition of a stellar surface map with a transiting planet into its preimage and null space components, 
along with their corresponding contributions to the light curve. Crucially, we observe that the planetary trajectory covers the visible part in the preimage 
component but absent in the null space. The transit path effectively "carves out" a strip on the stellar surface in nullspace but that is 
fully contained within the preimage.

\begin{figure*}[hbt!]
    \includegraphics[width=\textwidth]{figures/nullspace-preimage.png}
        \caption{The breakdown of a surface map into its constituent components: the original map (left), its preimage (center), and null space (right). 
        Accompanying each set is the corresponding impact on the star's rotational light curve. The preimage represents surface patterns that 
        directly influence the observed light curve, while the null space encompasses patterns that have no effect on the flux measurements. 
        It's noteworthy that a significant portion of the surface features fall within the null space, meaning they do not contribute to the light curve 
        that we can observe and measure. The blck line shows the planetary trajectory, so what is important is that the preimage contains the full
        information under planetary trajectory, while the nullspace is missing it.}
        \label{fig:nullspace}
    \end{figure*}
% In this work, we develop a Gaussian process framework to model the surfaces of exoplanet-host stars based on time-series photometry. We utilize 
% expressive kernel function from \cite{Luger2021b}. We first use this model on a simulated data to test the model, then we apply the model to an active K-dwarf
% HAT-P-11.

% The planet-hosting star HAT-P-11 has been extensively studied for its star spots. HAT-P-11 is orbited by a Neptune-sized exoplanet, HAT-P-11b, 
% which was discovered in 2008 via the transit method \citep{Bakos2010,Deming2011}. The transiting nature of HAT-P-11b has allowed for detailed 
% characterization of the star spot activity on HAT-P-11, as star spots that are occulted during the planet's transit cause detectable brightening in 
% the transit light curve \citep{Morris2017}.

%
% \section{The Data}
% The data is collected by the \emph{Kepler} mission and we extracted it using
% \texttt{lightkurve}, a Python package for Kepler and TESS data analysis \citep{lightkurve}.
% %
% \begin{figure}[ht!]
%     \script{TransitFitsWithStarry.py}
%     \begin{centering}
%         \includegraphics[width=\linewidth]{figures/TransitFitsWithStarry.pdf}
%         \caption{
%             Transit fits with starry -- no GPs.
%         }
%         \label{fig:TransitFitsStarry}
%     \end{centering}
% \end{figure}
%
\section{Hierarchical Bayesian model}
\label{sec:model}
In this section, we describe our Gaussian Process (GP) model and the likelihood calculation process used to estimate the model parameters.
This essentially means that we approximate the likelihood 
as a multidimensional Gaussian distribution characterized by a mean $\pmb{\mu}$ and covariance $\pmb{B}$ (for a comprehensive 
derivation of the Gaussian Process, refer to \cite{Luger2021b}):
%
\begin{align}
    \mathcal{L}\left(i_\star, P_\star, \mathbf{u}, \pmb{\theta}_\bullet\right) \sim
    \mathcal{N}\Big(
    \pmb{\mu}\left(i_\star, P_\star, \mathbf{u}, \pmb{\theta}_\bullet\right),
    \,
    \pmb{B}\left(i_\star, P_\star, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \Big).
\end{align}
%
To infer the physical properties of the starposts, we want to obtain an underlying distribution of the physics conditioned on the provided data, 
which is known as the posterior distribution of the model.
% \subsection{Model for non-evolvng surfaces} \label{sec:nonevolmodel} 
We want to solve for a large set of parameters that includes the GP hyperparameters, the star's and the planet's orbital parameters, respectively. 
\begin{linenomath}\begin{align}
    \label{eq:largetheta}
    \pmb{\Theta}
     & =
    \left(
    \theta_\bullet
    \,\,\,
    \theta_\star
    \,\,\,
    \theta_p
    \right)^\top
    \quad,
\end{align}\end{linenomath}

@@

We represent the GP hyperparameters as \emph{physically interesting} set of parameters $\pmb{\theta}_\bullet$ \citep{Luger2021b}:
%
\begin{linenomath}\begin{align}
        \label{eq:thetaspot}
        \pmb{\theta}_\bullet
         & =
        \left(
        \mathbb{n}
        \,\,\,
        \mathbb{c}
        \,\,\,
        \mathbb{a}
        \,\,\,
        \mathbb{b}
        \,\,\,
        \mathbb{r}
        \right)^\top
        \quad,
    \end{align}\end{linenomath}
%
where $\mathbb{n}$ is the number of starspots, $\mathbb{c}$ is their contrast (defined as the intensity difference between the spot and the 
background intensity, as a fraction of the background intensity), and $\mathbb{r}$ is the radius
of the spots. $\mathbb{a}$ and $\mathbb{b}$ are the normalized parameters of the Beta distribution in $\cos\phi$, which is the probability density function (PDF) 
for the latitude $\phi$ of the spots. These parameters have a one-to-one correspondence to the mode $\mu_\phi$ and standard deviation $\sigma_\phi$ of the 
distribution in $\phi$, allowing for a concise representation of the latitudinal distribution characteristics.
The Beta distribution in $\cos\phi$ which has hyperparameters $\alpha$ and $\beta$, and the PDF given by
%
\begin{align}
    \label{eq:cosphi-pdf}
    p \big(\cos\phi \, \big| \, \alpha, \beta \big)
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    (\cos\phi)^{\alpha - 1}
    (1 - \cos\phi)^{\beta - 1}
    \quad,
\end{align}
%
where $\Gamma$ is the Gamma function. The $\alpha$ and $\beta$ are derived from the normalized parameters 
%
\begin{align}
    \label{eq:beta2gauss}
    \alpha & = \exp\left({K_{00} + (\ln\frac{1}{2})\mathbb{a}}\right)
    \nonumber                                                 \\
    \beta  & = \exp\left({\ln\frac{1}{2} + (10 - \ln\frac{1}{2})\mathbb{b}}\right)
    \quad,
\end{align}
%
with inverse transform
%
\begin{align}
    \label{eq:gauss2beta}
    \mathbb{a} & \equiv \frac{\ln\alpha}{\ln\frac{1}{2}}
    \nonumber                                             \\[0.5em]
    \mathbb{b} & \equiv \frac{\ln\beta - \ln\frac{1}{2}}{10 -\ln\frac{1}{2}}
\end{align}
%
The parameters $\mathbb{a}$ and $\mathbb{b}$ are both constrained to values between 0 and 1, making them convenient for sampling during the inference process. 
However, $\mathbb{a}$ and $\mathbb{b}$ do not have a straightforward relationship with physically meaningful quantities. In many situations, it is preferable to parametrize 
the latitude distribution using two parameters: $\mu_\phi$, which controls the central latitude, and $\sigma_\phi$, which governs the dispersion of the 
spots' latitudes. Here, both the mean $\mu_\phi$ and variance $\sigma^2_\phi$ can be derived from the Beta distribution as
%
\begin{align}
    \label{eq:mean_beta}
    \mu_\phi
     & =
    \dfrac{\alpha}{\alpha+\beta}
    \quad,
\end{align}
%
\begin{align}
    \label{eq:var_beta}
    \sigma^2_\phi
     & =
    \dfrac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}
    \quad,
\end{align}
%

We assume that the prior over $\mathbb{f}_{true}$ follows a multivariate Gaussian distribution, with a mean vector of zeros and a covariance 
matrix $\pmb{B}$. We use the quasi-periodic kernel to define the covariance matrix $\pmb{B}$, which is defined by \texttt{StarryProcess}.
We assume that the observations $\mathbb{f}_{obs}$ are corrupted by additive Gaussian noise, such that:
\begin{equation}
    \mathbb{f}_{obs} = \mathbb{f}_{true} + \epsilon
\end{equation}

@@

\subsection{The issue of the normalization of light curves and units}
Before going on discussing the experiments produced for this paper, we first need to remind the reader of a subtlety of the flux normalization when we are given
the raw light curves from telescopes. The problem and the ways to tackle it were described in \cite{Luger2021a} and \cite{Luger2021b}.

@@

In our experiments, we noticed that dividing by a constant factor \emph{and} adding an additive baseline messes up the units due to the 
intricate transition between the spherical harmonics and flux unit bases. To address this issue, we opt for an alternative approach, wherein we 
impose a constraint upon the constant terms within the map, which correspond to a featureless star. Specifically, we achieve this by setting the prior distribution 
over these constant terms to be entirely uniform, thereby setting the first row of the precision matrix to zero, which is us telling the model that
\textit{we don't know what the baseline is}.

Let $\mathbf{K}$ be the precision matrix, which is the inverse of the covariance matrix $\mathbf{B}$. Then the precision matrix looks as
%
\begin{linenomath}\begin{align}
    \mathbf{K} = 
    \begin{bmatrix}
    0 & 0 & \cdots & 0 \\
    0 & k_{22} & \cdots & k_{2n} \\
    \vdots & \vdots & \ddots & \vdots \\
    0 & k_{n2} & \cdots & k_{nn}
    \end{bmatrix}
\end{align}\end{linenomath}
%
where $k_{ij}$ represents the elements of the precision matrix $\mathbf{K}$.

By setting the first row of the precision matrix to zero, we effectively impose a uniform prior distribution over the constant terms in the map, thereby 
addressing the issue of unit mismatch caused by dividing by a constant factor and adding an additive baseline.

@@

\subsection{Stellar inclination and obliquity}
\label{sec:obl-inc}
In the Sectrion \ref{sec:results-synthetic}, we hope to convince the reader that there is a a good amount of information about stars that we can learn 
from disk-integrated photometric measurements. In particular, Figure \ref{fig:inclination} illustrates the profound impact of stellar inclination on observed light curves. 
As the inclination angle varies from pole-on ($0^\circ$) to equator-on ($90^\circ$) views, we observe significant changes in both the amplitude and 
morphology of the light curve modulations. At low inclinations, when we view the star more pole-on, the light curves exhibit smaller amplitude variations. 
This is because the spots remain invisible throughout the stellar rotation, leading to more subtle flux changes. As the inclination increases towards an 
equator-on perspective, the amplitude of the variations typically increases, and the shape of the light curve becomes more pronounced. These changes occur because 
at higher inclinations, spots periodically disappear from view as the star rotates, causing more dramatic swings in the observed flux. 
The effect is particularly noticeable for spots at mid-latitudes, which can produce sharp dips in the light curve as they rotate into and out of view. 
It's important to note that while inclination significantly affects the observed light curve, it does not alter the underlying rotation period of 
the star or the intrinsic properties of the spots themselves.

While variations in stellar obliquity can have significant implications for planetary systems, their effects on the overall stellar light curve 
are often not noticeable. The global light curve is primarily influenced by the star's rotation and the distribution of spots on its surface, 
with obliquity playing a less prominent role. However, the impact of stellar obliquity becomes much more apparent during planetary transits, 
particularly through the phenomenon of spot-crossing events. When a planet transits its host star, it can occult starspots along its path. 
These spot-crossing events manifest as brief increases in brightness during the transit, as the planet temporarily blocks a darker region of the stellar 
surface. The frequency, timing, and duration of these spot-crossing events are strongly influenced by the alignment between the stellar spin axis and 
the planet's orbital plane - in other words, the stellar obliquity. Figure \ref{fig:obliquity} shows three consecutive transit light curves for 
different stellar obliquities. Each row in the figure represents a different obliquity scenario, with obliquity increasing from left to right. 
The rows display three consecutive transits for each scenario. 
In systems with low obliquity, where the stellar equator is aligned with the 
planet's orbit, spot-crossing events tend to occur more regularly and predictably. This is evident in the first column of Figure \ref{fig:obliquity}, where the spot-crossing patterns show a more consistent appearance across consecutive transits. 
Conversely, in systems with high obliquity, the pattern of spot-crossings can be more erratic and less frequent. This is because the planet's transit path across 
the stellar disk intersects a different range of stellar latitudes, potentially missing spots concentrated at certain latitudes. The $90^\circ$ column of 
Figure \ref{fig:obliquity} demonstrates this, showing more varied and less predictable spot-crossing patterns between consecutive transits. 
Therefore, while the full-disk light curve may not clearly reveal changes in stellar obliquity, careful analysis of in-transit light curves, 
particularly focusing on spot-crossing events, can provide valuable insights into the star's orientation relative to the planetary orbit.

\begin{figure*}[hbt!]
    % \script{experiment-1-true-map.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/inclinations.pdf}
        \caption{
            Variation in stellar light curves as a function of inclination angle. Each curve displays the photometric time series (light curve) for a 
            hypothetical spotted star observed at different inclination angles. The x-axis represents time in days, while the y-axis shows the relative flux. 
            Inclination angles range from $0^\circ$ (pole-on view) to $90^\circ$ (equator-on view). Note how the amplitude and shape of the light 
            curve modulations change significantly with inclination, demonstrating the importance of this parameter in interpreting stellar variability. 
            The star's rotation period and spot configuration remain constant across all light curves.
        }
        \label{fig:inclination}
    \end{centering}
\end{figure*}

\begin{figure*}[hbt!]
    % \script{experiment-1-true-map.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/obliquities.pdf}
        \caption{
            Impact of stellar obliquity on consecutive planetary transit light curves. Each column represents a different stellar obliquity scenario, 
            with obliquity increasing from left to right. The rows show three consecutive transits for each scenario. The x-axis represents time relative to 
            the transit center, while the y-axis shows the normalized flux. Note the varying patterns of in-transit brightness fluctuations 
            (spot-crossing events) across different obliquities and between consecutive transits. These differences arise from the changing 
            geometry between the planet's transit path and the distribution of starspots at different stellar latitudes. 
            The consistent transit depth and duration across all panels indicate that other system parameters remain constant, 
            isolating the effect of stellar obliquity.
        }
        \label{fig:obliquity}
    \end{centering}
\end{figure*}

While the analysis of spot-crossing events in transit light curves provides a powerful tool for constraining stellar inclination and obliquity, 
there exists an important degeneracy in these measurements. Specifically, a system with stellar inclination $i_\star$ and obliquity $-\lambda_\star$ 
will produce identical light curves to a system with inclination $180-i_\star$ and obliquity $\lambda_\star$, but with a surface map flipped upside down.

This degeneracy arises from the symmetry in the geometry of the star-planet system. In both scenarios, the relative orientation between the stellar 
rotation axis and the planet's orbital plane remains the same, merely mirrored. As a result, the pattern of spot crossings and their evolution over 
multiple transits will be indistinguishable between these two configurations.

For example, a star with an inclination of $80^\circ$ and an obliquity of $-30^\circ$ will present the same transit light curves 
as a star with an inclination of $100^\circ$ and an obliquity of $30^\circ$ with a flipped map (see Figure \ref{fig:inc-obliquity-degeneracy}). 
In both cases, the angle between the stellar spin axis and the planet's orbital axis is identical, just reflected about the plane of the sky.
This degeneracy highlights a limitation in our ability to uniquely determine stellar geometry solely from transit light curve analysis. 
While we can constrain the relative alignment between the stellar spin and planetary orbit with high precision, we cannot distinguish between these 
mirrored configurations without additional information or observations.

\begin{figure}[hbt!]
    % \script{experiment-1-true-map.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/inc-obl-degeneracy.png}
        \caption{
            Illustration of the degeneracy between stellar inclination ($i_\star$) and obliquity ($\lambda_\star$) in transit light curves. 
            The left panel shows a star-planet system with inclination $i_\star$ and obliquity $-\lambda_\star$, while the right panel 
            depicts a system with inclination $180-i_\star$ and obliquity $\lambda_\star$ and with a flipped map. 
            Despite the different geometries, both configurations result in identical transit light curves, shown in the central plot. 
            The x-axis represents time relative to the transit center, and the y-axis shows normalized flux. 
            This degeneracy demonstrates that while transit light curves can constrain the relative alignment between stellar spin and planetary orbit, 
            they cannot distinguish between these mirrored configurations without additional information. 
            The angles are represented with $x$, $y$, and $z$ vectors.
        }
        \label{fig:inc-obliquity-degeneracy}
    \end{centering}
\end{figure}

Given the degeneracy between stellar inclination and obliquity, a naive sampling approach could lead to unnecessary computational costs and potential 
convergence issues. To address this, we can leverage the fact that inclination and obliquity can be represented as vectors in three-dimensional 
space (see Figure \ref{fig:inc-obliquity-degeneracy}). This allows us to implement a more efficient sampling strategy that breaks the degeneracy 
while maintaining physical consistency.
